% Last Updated: 09/10/2024
\documentclass[a4paper, twoside]{article}
% \documentclass[a4paper, twoside]{amsart}
\input{/Users/wenc/Documents/Math/universal-pre}

\begin{document}
\ifmain
    \title{\tbf{Title}}
    \author{Cong Wen, 09/24}
    \makeatletter
    \@ifclassloaded{amsart}{
        \address{X University}
        \email{sparkwenc@gmail.com}
    }{\date{}}
    \makeatother

    % \maketitle
    % \begin{abstract}
    %     Notes\footnote{Last updated on January 01, 2000}.
    % \end{abstract}

    \tableofcontents
\fi
% ----------------------------------------------------------------------------------------------------


\ssc{Beauville-Bogomolov decomposition}

Beauville-Bogomolov decomposition\cite{Bogomolov1974,Beauville1983Kahler} is the structure theorem for complex compact K{\"a}hler manifolds with torsion first Chern class, settled in 1980s. To learn, we collect statements from \href{https://math.unice.fr/~beauvill/conf/decomp.pdf}{Beauville's slides}.


\begin{thm}
	Let $M$ be a complex compact K{\"a}hler manifold with $c_{1}(M)=0\in \mathrm{H}_{}^{2}(M;\bR)$, then there exists a finite {\'e}tale $\oT{M}\ar M$ such that
	\[
		\oT{M}=T\tms\prod_{i}X_{i}\tms\prod_{j} Y_{j},
	\]
	where $T$ is a \tbf{complex torus}, $X_{i}$ are \tbf{Calabi-Yau} manifolds, $Y_{j}$ are \tbf{hyper-K{\"a}hler} manifolds.
\end{thm}


The decomposition theorem is a direct consequence of Yau's theorem\cite{Yau1978}, the proof of the decomposition can be stated in two main lemmas.


\begin{lem}
	Let $M$ be a complex compact K{\"a}hler manifold with $c_{1}(M)=0\in \mathrm{H}_{}^{2}(M;\bR)$, then there exists a finite {\'e}tale $T\tms N\ar M$, where $T$ is a complex torus, $N$ a compact K{\"a}hler manifold with $\pi_{1}(N)=0, K_{N}\cong\cO_{N}$.
\end{lem}

\begin{prf}[Sketch of the proof]
	Yau's theorem says $M$ is Ricci-flat. Cheeger-Gromoll's splitting says $M\cong(\bC^{k}\tms N)/\Gma$, with $\pi_{1}(N)=0, K_{N}\cong\cO_{N}$ and $\Gma\sbs\Aut(\bC^{k})\tms\Aut(N)$. Bieberbach's theorem says there exists a finite index subgroup $U<\Gma$, acting trivially on $N$ and by translations on $\bC^{k}$.

	Therefore, $(\bC^{k}\tms N)/U\cong T\tms N\ar M$ is finite {\'e}tale.
\end{prf}

\begin{lem}
	Let $N$ be a compact K{\"a}hler manifold with $\pi_{1}(N)=0, K_{N}\cong\cO_{N}$, then
	\[
		N\cong\prod_{i}X_{i}\tms\prod_{j}Y_{j},
	\]
	where $X_{i}$ are Calabi-Yau, $Y_{j}$ are irreducible symplectic.
\end{lem}

\begin{prf}[Sketch of the proof]
	Yau's theorem says the holonomy group of $N$ is contained in $\SU$. de Rham's decomposition together with Berger's classification says
	\[
		N\cong\prod_{i}X_{i}\tms\prod_{j}Y_{j},
	\]
	where the holonomy groups of $X_{i}$ and $Y_{j}$ are $\SU$ and $\Sp$ respectively. Bochner's principle computes
	\[
		\mathrm{H}_{}^{0}(X,\Oga_{X}\xB)=\bC\op\bC\oga, \mathrm{H}_{}^{0}(Y,\Oga_{Y}\xB)=\bC[\sgm],
	\]
	where the former is Calabi-Yau, the latter is hyper-K{\"a}hler.
\end{prf}


\sct{hyper-K{\"a}hler geometry}

We collect many facts about hyper-K{\"a}hler manifolds, mainly from Huybrechts' paper\cite{Huybrechts1999}, as well as many statements from the notes of \href{https://people.bath.ac.uk/masgks/Papers/kinosaki19.pdf}{Sankaran} and \href{https://www.math.ens.psl.eu/~debarre/HKmanifolds.pdf}{Debarre}.


Any K3 surface admits an interesting $0$-cycle $c_{X}\in\CH{0}{}{(X)}$:
\begin{thm}[{\cite[Theorem~1]{BV2004}}]
	Let $X$ be a K3 surface, then there exists $c_{X}\in\CH{0}{}{(X)}$ such that
	\begin{enr}
		\item Any point on a (possibly singular) rational curve in $X$ has class $c_{X}\in\CH{0}{}{(X)}$.
		\item The image of the intersection product $\Pic(X)\ot\Pic(X)\ar\CH{0}{}{(X)}$ is contained in $\bZ c_{X}$.
		\item $c_{2}(X)=24c_{X}\in\CH{0}{}{(X)}$.
	\end{enr}
\end{thm}


\sct{DHMV2022}


The paper proved certain hyper-K{\"a}hler fourfold is of K3\ts{[2]}  deformation type. The key is to prove the hyper-K{\"a}hler SYZ conjecture conditionally.

To learn, we collect key parts of the paper as follows. Notations conform with those in loc. cit.


\begin{cnj}
	Any hyper-K{\"a}hler manifold can be deformed into one with a Lagrangian fibration.
\end{cnj}

\begin{cnj}
	Let $X$ be a hyper-K{\"a}hler manifold of dimension $2n$, then any non-trivial nef line bundle $L$ on $X$ satisfying $\int_{X}c_{1}(L)^{2n}=0$ is semi-ample.
\end{cnj}





% ----------------------------------------------------------------------------------------------------
\ifmain
    \bibliographystyle{amsalpha}
    \bibliography{/Users/wenc/Documents/Math/universal-ref}
\fi
\end{document}
